\chapter*{Acknowledgments}

%{\small{The acknowledgment is intentionally excluded from this draft and will be added to the final version.}}

	
Throughout the last decade, Martin Frank has been my teacher and mentor. 
 With his rare ability to combine enthusiasm, rigor, and ease, he sparked my interest in mathematics the moment I first entered his \textit{Introductory Mathematics} lecture at RWTH Aachen University in 2009. Paraphrasing one of his favorite analogies,   
he has been the shepherd that guided me to the fruitful, green grass that is the field of kinetic theory. 
A uniquely approachable advisor, his constant guidance and the many insightful discussions have been integral to my academic and personal development.
From the beginning, he introduced me to the scientific community and taught me to communicate concisely and clearly towards peers and non-peers alike. 
For all this and much more, I want to thank Martin Frank and express my gratefulness.
% \textit{Thank you}.


Edward Larsen is one of the people that Martin Frank introduced to me. Discussing non-classical particle transport in front of a blackboard back in Aachen is one of the highlights of my years as a Ph.D. student. 
His experience in and deep understanding of transport theory is all the more impressive, given that it was always paired with a welcoming openness towards input from everyone involved in a discussion---even from inexperienced students like me. Our many (written) conversations made me a more knowledgeable person and a better writer, too. I am immensely grateful for these experiences.
Additionally, his many remarks helped to improve the quality of my work. 

I want to thank Gudrun Thäter for her fruitful input, for reading and improving my thesis, and for being part of my Ph.D. committee. Her help made this a better, more understandable thesis.


My colleagues, both in Aachen and Karlsruhe, have been a constant source of help and support. 
The many fruitful discussions broadened my understanding of mathematics and science in general. At the same time, I was fortunate to call them my friends. In particular, I want to thank Kerstin Küpper and Birte Schmidtmann for making the time at MathCCES unforgettable;  Maren Hattebuhr, Sarah Schönbrodt, Maike Sube, and Kirsten Wohak for teaching me the importance of making mathematics accessible to a diverse audience and future generations; and 
Jonas Kusch and Jannick Wolters for helping me overcome many challenging problems and always being there to answer questions about kinetic theory. I will miss our coffee breaks.

I also want to thank everyone I have ever run a single mile with. Be it with my friends from the Aachener TG, the LG Schanz, or the KIT SC, a solid run with good company has never failed to brighten my day. Additionally, Wolfgang Glöde taught me many lessons about mental and physical strength. I wouldn't be where I am today without the balance that running---and the many, many people I met along the way---provided.

My family's never-ending support and love helped me overcome many stressful moments throughout my life. They have always been there for me, regardless of the paths I took or the decisions I made. I couldn't wish for better parents. 

Lastly, my deepest gratitude goes to Nayana Karkanematt Jagadish. Both her joyful nature  and motivating words have been invaluable.


%}









